---
title: "PS demo2: process a real Perturb-seq data"
output: html_notebook
---

# Introduction

In this demo, we will process one BeeSTING-seq data from [Morris et al. Science 2023](https://www.science.org/doi/10.1126/science.adh7699). This is a very interesting base editing Perturb-seq dataset, but we will primarily use this dataset as an example to show how to run PS from the very beginning.

# Preprocessing

Load the R libraries


```{r}

library(Seurat)
library(ggplot2)
library(patchwork)
library(scales)
library(dplyr)

```

If you have not installed scMAGeCK yet, use the following command to install:


```{r}
if(FALSE){
  library(devtools)
  install_github('weililab/scMAGeCK') 
}
```

Once the installation is complete, load scMAGeCK:
```{r}
library(scMAGeCK)
```

Let's first read the expression data


```{r}

exp_mat_GDO=ReadMtx(mtx='counts/GSM7108136_BeeSTINGseq_GDO-A_matrix.mtx.gz',
                    cells='counts/GSM7108136_BeeSTINGseq_GDO-A_barcodes.tsv.gz',
                    features='counts/GSM7108136_BeeSTINGseq_GDO-A_features.tsv.gz')

sobj=CreateSeuratObject(counts=exp_mat_GDO[1:36601,])
sobj[['CRISPR']]=CreateAssayObject(counts=exp_mat_GDO[36602:36939,])

```


The rest is the standard Seurat preprocessing steps. See [Seurat tutorial](https://satijalab.org/seurat/articles/get_started_v5_new) for more details.

```{r}

sobj[["percent.mt"]] <- PercentageFeatureSet(sobj, pattern = "^MT-")
VlnPlot(sobj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

plot1 <- FeatureScatter(sobj, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(sobj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2

# sobj <- subset(sobj, subset = nFeature_RNA > 200 & nFeature_RNA < 4500 & percent.mt < 10)

sobj <- NormalizeData(object = sobj) %>% FindVariableFeatures() %>% ScaleData()

sobj <- RunPCA(object = sobj)

sobj <- RunUMAP(object = sobj, dims = 1:10)


sobj <- FindNeighbors(sobj, dims = 1:10)
sobj <- FindClusters(sobj, resolution = 0.5)
DimPlot(sobj, reduction = "umap")

```


# Calculate PS scores 

Now we start with PS calculation. 

## Step 1: prepare barcode files

First, we need to extract the barcode file which will be necessary for running PS:

```{r}


bc_frame=guidematrix_to_triplet(sobj[['CRISPR']]@counts,sobj)
bc_frame[,'sgrna']=bc_frame[,'barcode']
bc_frame[,'gene']=sub('-[0-9]+$','',bc_frame[,'barcode'])

# filter barcode file to only keep guides with >1 read counts

bc_frame <- bc_frame[bc_frame$read_count>1,]

```

## Step 2: pre-process guide expression matrix



```{r}

DefaultAssay(sobj)='CRISPR'

sobj <- pre_processRDS(bc_frame, sobj)

FeaturePlot(sobj,features=c('NTC-GV2','CD55-Q86X'))
```

## Step 3: run PS method

Use *scmageck_eff_estimate* function to calculate PS score. For more information, use *?scmageck_eff_estimate* 

```{r}
DefaultAssay(sobj)<-'RNA'

# test CD55

eff_object <- scmageck_eff_estimate(sobj, bc_frame, perturb_gene=c('SNP-36'), 
                                    non_target_ctrl = 'NTC-GV2',subset_rds = T, lambda = 0, target_gene_max = 100)

eff_estimat=eff_object$eff_matrix
rds_subset=eff_object$rds  

```


Now you can visualize the PS scores

```{r}
FeaturePlot(rds_subset,features='SNP.36_eff')
FeaturePlot(rds_subset,features = 'APPBP2')

```

We can further perform additional analysis, for example, to see whether CD55 mutations affect CD55 expression:

```{r}

data_f=FetchData(rds_subset,vars = c('sgrna','gene','APPBP2','SNP.36_eff'))

library(ggplot2)

data_f$PS=ifelse(data_f$gene=='NTC-GV2','NTC',ifelse(data_f$SNP.36_eff > 0.5, 'PS_High','PS_Low'))
#data_f$PS=ifelse(data_f$CD55.Q86X_eff > 0.5, 'PS_High','PS_Low')

p<-ggplot(data_f,aes(x=PS,y=APPBP2)) + 
  geom_violin() + 
  theme_bw()

print(p)


```

An ECDF plot can be shown for different groups:

```{r}


plot(ecdf(data_f$APPBP2[data_f$gene=='NTC-GV2']),ylim=c(0.9,1),xlab='APPBP2 expression',ylab='Fraction')
plot(ecdf(data_f$APPBP2[data_f$SNP.36_eff<0.2]),add=T,col='blue')
plot(ecdf(data_f$APPBP2[data_f$SNP.36_eff>0.2]),add=T,col='red')

legend('bottomright',c('NT','PS low','PS high'),col=c('black','blue','red'),lwd=1,pch=20)

```


