---
title: "Pancreatic Differentiation scRNA-seq"
author: 
- name: "Bicna Song"
- affiliation: "Children's National Hospital"
- name: "Wei Li"
- affiliation: "Children's National Hospital"
date: "2023-12-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup the Seurat Object

```{r}
library(Seurat)
library(scMAGeCK)
library(hdf5r)
library(ggplot2)
```

```{r}
rds <- readRDS("10clones_seurat.rds")
```

```{r}
DimPlot(rds,reduction = "umap")
```

```{r}
bc_frame <- rds@meta.data[,c("sgrna","gene","nCount_sgRNA")]

bc_frame[,"cell"] <- rownames(bc_frame)
bc_frame[,"barcode"] <- bc_frame[,"sgrna"]

colnames(bc_frame) <- c("sgrna","gene","read_count","cell","barcode")

bc_frame <- bc_frame[,c("cell","barcode","sgrna","gene","read_count")]

bc_frame[,"umi_count"] <- bc_frame[,"read_count"]
bc_frame <- bc_frame[!is.na(bc_frame$gene),]

bc_frame$gene <- sub("_","-",bc_frame$gene)
head(bc_frame)
```

```{r}
write.table(bc_frame, file = "10clones_seurat_scmageck_processed.frame.txt", row.names = F, quote = F, sep = "\t")
```

```{r}
rds <- pre_processRDS(bc_frame,rds)
```

```{r}
table(rds$gene)
```

```{r}
targetgenelist <- c("43-HHEX", "44-HHEX", "45-HHEX", "51-FOXA1", "53-FOXA1/2", "57-OTUD5", "58-OTUD5", "61-CCDC6", "62-CCDC6")

targetgenelist_geneid <- c("HHEX","HHEX","HHEX","FOXA1","FOXA2","OTUD5","OTUD5","CCDC6","CCDC6")

negative_ctrl_gene <- "47-WT" 
```

```{r}
eff_obj <- scmageck_eff_estimate(rds,bc_frame, targetgenelist, negative_ctrl_gene,
                                 scale_factor = 6, assay_for_cor = "RNA",
                                 perturb_gene_exp_id_list = targetgenelist_geneid)
```

```{r}
eff_estimat <- eff_obj$eff_matrix
rds_subset <- eff_obj$rds
```

```{r}
for(pb_i in 1:length(targetgenelist)){
  pb <- targetgenelist[pb_i]
  
  p <- FeaturePlot(rds_subset, features = paste(pb, "eff", sep = "_"))
  print(p)
  
  pb_gene <- targetgenelist_geneid[pb_i]
  p <- FeaturePlot(rds_subset,features = pb_gene)
  print(p)
}
```

```{r}
if(F){
  saveRDS(rds_subset, file = "10clones_seurat_scmageck_processed.rds")
}else{
  rds_subset = readRDS("10clones_seurat_scmageck_processed.rds")
}
```

```{r}
if(F){
  saveRDS(eff_obj, file = "10clones_seurat_scmageck_processed_eff_object.rds")
}else{
  eff_obj = readRDS("10clones_seurat_scmageck_processed_eff_object.rds")
}
```

# Generating CCDC6 PS (Figure 5f, Figure S10a,b)

```{r}
if(F){
  mk1 <- FindMarkers(rds_subset, group.by = "gene", ident.1 = "61-CCDC6", ident.2 = "47-WT", logfc.threshold = 0.1)
  res_1 <- eff_obj$target_gene_search_result$`61-CCDC6`$deframe
  
  rds_c47 <- subset(rds_subset, seurat_clusters %in% c("4","7","10")) # LV/DUO
  mk1_47 <- FindMarkers(rds_c47, group.by = "gene", ident.1 = "61-CCDC6", ident.2 = "47-WT", logfc.threshold = 0.1)
  
  rds_c01513 <- subset(rds_subset, seurat_clusters %in% c("0","1","5","13")) # DE
  mk1_01513 <- FindMarkers(rds_c01513, group.by = "gene", ident.1 = "61-CCDC6", ident.2 = "47-WT", logfc.threshold = 0.1)
  
  rds_c6_PP <- subset(rds_subset,seurat_clusters %in% c("6")) # PP
  mk1_6_PP <- FindMarkers(rds_c6_PP, group.by = "gene", ident.1 = "61-CCDC6", ident.2 = "47-WT", logfc.threshold = 0.1)
  
  rds_c63812_PP <- subset(rds_subset, seurat_clusters %in% c("6","3","8","12")) # PP/PP in transition
  mk1_63812_PP <- FindMarkers(rds_c63812_PP, group.by = "gene", ident.1 = "61-CCDC6", ident.2 = "47-WT", logfc.threshold = 0.1)
  
  rds_c2_DE <- subset(rds_subset, seurat_clusters %in% c("2")) # DE in transition
  mk1_2_DE <- FindMarkers(rds_c2_DE, group.by = "gene", ident.1 = "61-CCDC6", ident.2 = "47-WT", logfc.threshold = 0.1)
  
  m1k_merge <- merge(mk1_47, mk1_01513, by = 0, suffixes = c("_c47","c_01513"), all = T)
  m1k_merge <- merge(m1k_merge, mk1, by.x = 1, by.y = 0, all = T)
  row.names(m1k_merge) = m1k_merge[,1]
  
  write.csv(m1k_merge, file = "table/61_ccdc6_wt_c47_c01513.csv", quote=F)
  write.csv(mk1_47, file = "table/61_ccdc6_wt_c47.csv")
  write.csv(mk1_01513, file = "table/61_ccdc6_wt_c01513.csv")
  write.csv(mk1_6_PP, file = "table/61_ccdc6_wt_c6_PP.csv")
  write.csv(mk1_63812_PP, file = "table/61_ccdc6_wt_c63812_PP.csv")
  write.csv(mk1_2_DE, file = "table/61_ccdc6_wt_c2_DE.csv")
  
}else{
  mk1_47 <- read.csv("table/61_ccdc6_wt_c47.csv", row.names = 1)
  mk1_01513 <- read.csv("table/61_ccdc6_wt_c01513.csv", row.names = 1)
  mk1_6_PP <- read.csv("table/61_ccdc6_wt_c6_PP.csv", row.names = 1)
  mk1_63812_PP <- read.csv("table/61_ccdc6_wt_c63812_PP.csv", row.names = 1)
  m1k_merge <- read.csv("table/61_ccdc6_wt_c47_c01513.csv", row.names = 1)
  mk1_2_DE <- read.csv("table/61_ccdc6_wt_c2_DE.csv", row.names = 1)
}
```

```{r}
rds_subset2 <- rds_subset

for(sx in grep("_eff", colnames(rds_subset2@meta.data), value = T)){
  rds_subset2@meta.data[,sx] = NULL
}
```

```{r}
targetgenelist <- c("61-CCDC6" )
targetgenelist_geneid <- c("CCDC6")
negative_ctrl_gene <- "47-WT" 

eff_obj2 <- scmageck_eff_estimate(rds_subset2, bc_frame, targetgenelist, negative_ctrl_gene,
                      scale_factor = 6,
                      assay_for_cor = "RNA",
                      perturb_gene_exp_id_list = targetgenelist_geneid,
                      background_correction = T)
```

```{r}
eff_estimat2 <- eff_obj2$eff_matrix
rds_subset2_rerun <- eff_obj2$rds
```

```{r}
for(pb_i in 1:length(targetgenelist)){
  pb <- targetgenelist[pb_i]
  
  p <- FeaturePlot(rds_subset2_rerun, features = paste(pb, "eff", sep="_"))
  print(p)
  
  pb_gene <- targetgenelist_geneid[pb_i]
  p <- FeaturePlot(rds_subset2_rerun, features = pb_gene)
  print(p)
}
```

## PP/PP in transition (clusters 6,3,8,12)

```{r}
target_gene_63812 <- rownames(mk1_63812_PP)[abs(mk1_63812_PP$avg_log2FC) > 0.25 & mk1_63812_PP$p_val_adj < 0.05]

eff_obj5 <- scmageck_eff_estimate(rds_subset2, bc_frame, targetgenelist, negative_ctrl_gene,
                                  perturb_target_gene = target_gene_63812,
                                  scale_factor = 6,
                                  assay_for_cor = "RNA",
                                  perturb_gene_exp_id_list = targetgenelist_geneid,
                                  background_correction = T)
```

## Figure 5f - PP/PP in transtion

```{r}
p <- FeaturePlot(eff_obj5$rds, features = grep("eff", colnames(eff_obj5$rds@meta.data), value = T)) + ggtitle("Pattern 1: PP/PP in transition")
print(p)
```

## DE (clusters 0,1,5,13)

```{r}
target_gene_01513 <- rownames(mk1_01513)[abs(mk1_01513$avg_log2FC) > 0.25 & mk1_01513$p_val_adj < 0.05]

eff_obj3 <- scmageck_eff_estimate(rds_subset2, bc_frame, targetgenelist, negative_ctrl_gene,
                                  perturb_target_gene = target_gene_01513,
                                  scale_factor=6,
                                  assay_for_cor = "RNA", 
                                  perturb_gene_exp_id_list = targetgenelist_geneid,
                                  background_correction = T)
```

## Figure 5f - DE

```{r}
p <- FeaturePlot(eff_obj3$rds, features = grep("eff", colnames(eff_obj3$rds@meta.data), value = T)) + ggtitle("Pattern 2: DE")
print(p)
```

## LV/DUO (clusters 4,7,10)

```{r}
target_gene_47 <- rownames(mk1_47)[abs(mk1_47$avg_log2FC) > 0.25 & mk1_47$p_val_adj < 0.05]

eff_obj4 <- scmageck_eff_estimate(rds_subset2, bc_frame, targetgenelist, negative_ctrl_gene,
                                  perturb_target_gene = target_gene_47,
                                  scale_factor = 6,
                                  assay_for_cor = "RNA",
                                  perturb_gene_exp_id_list = targetgenelist_geneid,
                                  background_correction = T)
```

## Figure S10a - LV/DUO

```{r}
p <- FeaturePlot(eff_obj4$rds, features = grep("eff", colnames(eff_obj4$rds@meta.data), value = T)) + ggtitle("Pattern 1: CCD6 PS score from LV/DUO")
print(p)
```

## DE in transition (cluster 2)

```{r}
target_gene_c2 <- rownames(mk1_2_DE)[abs(mk1_2_DE$avg_log2FC) > 0.25 & mk1_2_DE$p_val_adj < 0.05]

eff_obj6 <- scmageck_eff_estimate(rds_subset2, bc_frame, targetgenelist, negative_ctrl_gene,
                                  perturb_target_gene = target_gene_c2,
                                  scale_factor = 6,
                                  assay_for_cor = "RNA",
                                  perturb_gene_exp_id_list = targetgenelist_geneid,
                                  background_correction = T)
```

## Figure S10b - DE in transition

```{r}
p <- FeaturePlot(eff_obj6$rds, features = grep("eff", colnames(eff_obj6$rds@meta.data), value = T)) + ggtitle("Pattern 2: CCD6 PS score from DE in transition")
print(p)
```

# Generating HHEX PS (Figure 5c)

```{r}
rds_subset3 <- rds_subset

for(sx in grep("_eff", colnames(rds_subset3@meta.data), value = T)){
  rds_subset3@meta.data[,sx]=NULL
}
```

```{r}
targetgenelist=c("44-HHEX", "45-HHEX" )
targetgenelist_geneid=c("HHEX", "HHEX")
negative_ctrl_gene="47-WT" 

eff_obj7 <- scmageck_eff_estimate(rds_subset3, bc_frame, targetgenelist, negative_ctrl_gene,
                                  target_gene_min = 50,
                                  target_gene_max = 100,
                                  scale_factor = 6,
                                  assay_for_cor = "RNA",
                                  perturb_gene_exp_id_list = targetgenelist_geneid,
                                  background_correction = T)
```

```{r}
eff_estimat3 <- eff_obj7$eff_matrix
rds_subset3_rerun <- eff_obj7$rds
```

```{r}
for(pb_i in 1:length(targetgenelist)){
  pb <- targetgenelist[pb_i]
  p <- FeaturePlot(rds_subset3_rerun, features = paste(pb, "eff", sep = "_"))
  print(p)
  
  pb_gene <- targetgenelist_geneid[pb_i]
  p <- FeaturePlot(rds_subset3_rerun, features = pb_gene)
  print(p)
}
```

```{r}
if(F){
mkhhex_1 <- FindMarkers(rds_subset3_rerun, group.by = "gene", ident.1 = "45-HHEX", ident.2 = "47-WT", subset.ident = c("0","1","5","13"), logfc.threshold = 0.1) # DE
mkhhex_2 <- FindMarkers(rds_subset3_rerun, group.by = "gene", ident.1 = "45-HHEX", ident.2 = "47-WT", subset.ident = c("4","7","10"), logfc.threshold = 0.1) # LV/DUO
mkhhex_3 <- FindMarkers(rds_subset3_rerun, group.by = "gene", ident.1 = "45-HHEX", ident.2 = "47-WT", subset.ident = c("6"), logfc.threshold = 0.1) # PP
mkhhex_36812 <- FindMarkers(rds_subset3_rerun, group.by = "gene", ident.1 = "45-HHEX", ident.2 = "47-WT", subset.ident = c("3","6","8","12"), logfc.threshold = 0.1) # PP/PP in transition
mkhhex_3_all <- FindMarkers(rds_subset3, group.by = "gene", ident.1 = c("43-HHEX", "44-HHEX", "45-HHEX"), ident.2 = "47-WT", subset.ident = c("6"), logfc.threshold = 0.1) # PP
mkhhex_36812_all <- FindMarkers(rds_subset3, group.by = "gene", ident.1 = c("43-HHEX", "44-HHEX", "45-HHEX"), ident.2 = "47-WT", subset.ident = c("3","6","8","12"), logfc.threshold = 0.1) # PP/PP in transition

write.csv(mkhhex_1, file = "table/45hhex_vs_wt_c01513.csv",quote=F)
write.csv(mkhhex_2, file = "table/45hhex_vs_wt_c4710.csv",quote=F)
write.csv(mkhhex_3, file = "table/45hhex_vs_wt_c6.csv",quote=F)
write.csv(mkhhex_3_all, file = "table/allhhex_vs_wt_c6.csv",quote=F)
write.csv(mkhhex_36812, file = "table/45hhex_vs_wt_c36812.csv",quote=F)
write.csv(mkhhex_36812_all, file = "table/allhhex_vs_wt_c36812.csv",quote=F)

}else{
  mkhhex_1 <- read.csv("table/45hhex_vs_wt_c01513.csv")
  mkhhex_2 <- read.csv("table/45hhex_vs_wt_c4710.csv")
  mkhhex_3 <- read.csv("table/45hhex_vs_wt_c6.csv")
  mkhhex_3_all <- read.csv("table/allhhex_vs_wt_c6.csv")
  mkhhex_36812 <- read.csv("table/45hhex_vs_wt_c36812.csv")
  mkhhex_36812_all <- read.csv("table/allhhex_vs_wt_c36812.csv")
}
```

```{r}
hhex_cluster01513_target_gene <- mkhhex_1[mkhhex_1$p_val_adj < 0.05 & abs(mkhhex_1$avg_log2FC) > 0.5, 1]
hhex_cluster4710_target_gene <- mkhhex_2[mkhhex_2$p_val_adj < 0.25, 1] # very few diff exp genes 
hhex_cluster6_target_gene <- mkhhex_3_all[mkhhex_3_all$p_val_adj < 0.05, 1]
hhex_cluster6_target_gene <- mkhhex_3[mkhhex_3$p_val_adj < 0.05 & abs(mkhhex_3$avg_log2FC) > 0.75, 1]
hhex_cluster36812_target_gene <- mkhhex_36812[mkhhex_36812$p_val_adj<0.05 & abs(mkhhex_36812$avg_log2FC) > 0.7, 1]
```

```{r}
rds_subset4 <- rds_subset

for(sx in grep("_eff",colnames(rds_subset4@meta.data), value = T)){
  rds_subset4@meta.data[,sx] = NULL
}
```

## DE (clusters 0,1,5,13)

```{r}
targetgenelist <- c("43-HHEX", "44-HHEX", "45-HHEX")
targetgenelist_geneid <- c("HHEX", "HHEX", "HHEX")
negative_ctrl_gene<- "47-WT" 

eff_obj8 <- scmageck_eff_estimate(rds_subset4, bc_frame, targetgenelist, negative_ctrl_gene,
                                  perturb_target_gene = hhex_cluster01513_target_gene,
                                  scale_factor = 6,
                                  assay_for_cor = "RNA",
                                  perturb_gene_exp_id_list = targetgenelist_geneid,
                                  lambda = 0.0,
                                  background_correction = T)
```

```{r}
eff_estimat4 = eff_obj8$eff_matrix
rds_subset4_rerun = eff_obj8$rds

for(pb_i in 1:length(targetgenelist)){
  pb <- targetgenelist[pb_i]
  p <- FeaturePlot(rds_subset4_rerun, features = paste(pb, "eff", sep="_"))
  print(p)
  
  pb_gene <- targetgenelist_geneid[pb_i]
  p <- FeaturePlot(rds_subset4_rerun, features = pb_gene)
  print(p)
}
```

## PP/PP in transition (clusters 3,6,8,12)

```{r}
eff_obj9 <- scmageck_eff_estimate(rds_subset4, bc_frame, targetgenelist, negative_ctrl_gene,
                                  perturb_target_gene = hhex_cluster36812_target_gene,
                                  scale_factor = 6,
                                  assay_for_cor = "RNA",
                                  perturb_gene_exp_id_list = targetgenelist_geneid,
                                  lambda = 0.0,
                                  background_correction = T)
```

```{r}
eff_estimat5 = eff_obj9$eff_matrix
rds_subset4_rerun = eff_obj9$rds

for(pb_i in 1:length(targetgenelist)){
  pb <- targetgenelist[pb_i]
  p <- FeaturePlot(rds_subset4_rerun, features = paste(pb, "eff", sep="_"))
  print(p)
  
  pb_gene <- targetgenelist_geneid[pb_i]
  p <- FeaturePlot(rds_subset4_rerun, features = pb_gene)
  print(p)
}
```

## LV/DUO (clusters 4,7,10)

```{r}
eff_obj10 <- scmageck_eff_estimate(rds_subset4, bc_frame, targetgenelist, negative_ctrl_gene,
                                   perturb_target_gene = hhex_cluster4710_target_gene,
                                   scale_factor = 6,
                                   assay_for_cor = "RNA",
                                   perturb_gene_exp_id_list = targetgenelist_geneid,
                                   lambda = 0.0,
                                   background_correction = T)
```

```{r}
eff_estimat6 = eff_obj10$eff_matrix
rds_subset4_rerun = eff_obj10$rds

for(pb_i in 1:length(targetgenelist)){
  pb <- targetgenelist[pb_i]
  p <- FeaturePlot(rds_subset4_rerun, features = paste(pb, "eff", sep="_"))
  print(p)
  
  pb_gene <- targetgenelist_geneid[pb_i]
  p <- FeaturePlot(rds_subset4_rerun, features = pb_gene)
  print(p)
}
```

```{r}
rds_subset5 <- rds_subset
for(sx in grep("_eff", colnames(rds_subset5@meta.data), value = T)){
  rds_subset5@meta.data[,sx]=NULL
}

z <- rds_subset5$gene
z[z=="43-HHEX"] <- "HHEX"
z[z=="44-HHEX"] <- "HHEX"
z[z=="45-HHEX"] <- "HHEX"

rds_subset5 <- AddMetaData(rds_subset5, z, col.name = "gene")

bc_frame2 <- bc_frame
z <- bc_frame2$gene
z[z=="43-HHEX"] <- "HHEX"
z[z=="44-HHEX"] <- "HHEX"
z[z=="45-HHEX"] <- "HHEX"
bc_frame2$gene  <-  z
```

```{r}
targetgenelist <- c("HHEX")
targetgenelist_geneid <- c("HHEX")
negative_ctrl_gene <- "47-WT" 

eff_obj11 <- scmageck_eff_estimate(rds_subset5, bc_frame2, targetgenelist, negative_ctrl_gene,
                                   perturb_target_gene = hhex_cluster36812_target_gene,
                                   scale_factor = 6,
                                   assay_for_cor = "RNA",
                                   perturb_gene_exp_id_list = targetgenelist_geneid,
                                   background_correction = T,
                                   lambda = 0)
```

```{r}
eff_estimat7=eff_obj11$eff_matrix
rds_subset5_rerun=eff_obj11$rds
```

## Figure 5c - HHEX PS score

```{r}
p <- FeaturePlot(subset(rds_subset5_rerun, gene == "HHEX"), features = "HHEX_eff") + ggtitle("HHEX PS score")
print(p)
```

# Generating FOXA1 PS (Figures S9a,b,c)

```{r}
rds_subset6 <- rds_subset

for(sx in grep("_eff",colnames(rds_subset6@meta.data), value = T)){
  rds_subset6@meta.data[,sx]=NULL
}
```

```{r}
targetgenelist <- c("51-FOXA1", "53-FOXA1/2" )
targetgenelist_geneid <- c("FOXA1", "FOXA1")
negative_ctrl_gene <- "47-WT" 

eff_obj12 <- scmageck_eff_estimate(rds_subset6, bc_frame, targetgenelist, negative_ctrl_gene,
                                   scale_factor = 6,
                                   assay_for_cor = "RNA",
                                   perturb_gene_exp_id_list = targetgenelist_geneid,
                                   background_correction = T)
```

```{r}
eff_estimat8 <- eff_obj12$eff_matrix
rds_subset6_rerun <- eff_obj12$rds
```

## Figure S9a - FOXA1 PS score (clone 51)

```{r}
p <- FeaturePlot(eff_obj12$rds, features = "51-FOXA1_eff") + ggtitle("FOXA1 PS score (clone 51)")
print(p)
```

## Figure S9b - FOXA1 PS score (clone 53)

```{r}
p <- FeaturePlot(eff_obj12$rds, features = "53-FOXA1/2_eff") + ggtitle("FOXA1 PS score (clone 53)")
print(p)
```

## Figure S9c - FOXA1 expression (all clones)

```{r}
p <- FeaturePlot(rds_subset6_rerun, features = "FOXA1") + ggtitle("FOXA1 expression (all clones)")
print(p)
```

```{r}
session_info()
```

